
## Load Packages

```{r}
options(scipen=999)
library(haven)
library(dplyr)
library(tidycensus)
library(janitor)
library(purrr)
library(lehdr)
library(sf)
library(tigris)
library(ggplot2)
library(scales)
library(tidyr) 
```

## Download Work Area Characteristics Data for Texas for each census tract

```{r}
wac  <- grab_lodes(state = "tx", 
                    year = c(2002, 2012, 2022), 
                    version = "LODES8", 
                    lodes_type = "wac", 
                    job_type = "JT00",
                    segment = "S000", 
                    state_part = "main", 
                    agg_geo = "tract")

tail(wac)  
```

# List the industries of interest by ID

```{r}
industryvars <- c("C000",  # Total jobs
                   "CNS01", # Agriculture, Forestry, Fishing and Hunting
                   "CNS02", # Mining, Quarrying, and Oil and Gas Extraction
                   "CNS03", # Utilities
                   "CNS04", # Construction
                   "CNS05", # Manufacturing
                   "CNS06", # Wholesale Trade
                   "CNS07", # Retail Trade
                   "CNS08", # Transportation and Warehousing
                   "CNS09", # Information
                   "CNS10", # Finance and Insurance
                   "CNS11", # Real Estate and Rental and Leasing
                   "CNS12", # Professional, Scientific, and Technical Services
                   "CNS13", # Management of Companies and Enterprises
                   "CNS14", # Admin & Support, Waste Mgmt & Remediation
                   "CNS15", # Educational Services
                   "CNS16", # Health Care and Social Assistance
                   "CNS17", # Arts, Entertainment, and Recreation
                   "CNS18", # Accommodation and Food Services
                   "CNS19", # Other Services (excl. Public Admin)
                   "CNS20"  # Public Administration
)
```


```{r}
industry_labels <- tribble(
  ~industry_code, ~Industry Name ,
  "C000", "Total Jobs",
  "CNS01", "Agriculture, Forestry, Fishing and Hunting",
  "CNS02", "Mining, Quarrying, and Oil and Gas Extraction",
  "CNS03", "Utilities",
  "CNS04", "Construction",
  "CNS05", "Manufacturing",
  "CNS06", "Wholesale Trade",
  "CNS07", "Retail Trade",
  "CNS08", "Transportation and Warehousing",
  "CNS09", "Information",
  "CNS10", "Finance and Insurance",
  "CNS11", "Real Estate and Rental and Leasing",
  "CNS12", "Professional, Scientific, and Technical Services",
  "CNS13", "Management of Companies and Enterprises",
  "CNS14", "Administrative and Support, Waste Management and Remediation",
  "CNS15", "Educational Services",
  "CNS16", "Health Care and Social Assistance",
  "CNS17", "Arts, Entertainment, and Recreation",
  "CNS18", "Accommodation and Food Services",
  "CNS19", "Other Services (excluding Public Administration)",
  "CNS20", "Public Administration"
)

```


# Get the geographic boundaries of the state of Texas: State and Places

```{r}
options(tigris_use_cache = TRUE)

tractstx <- tracts(state = "TX", year = 2020)
placestx <- places(state = "TX", year = 2020)
arlington <- placestx %>% filter(NAME == "Arlington")

```


# Keep only tracts whose centroids are inside Arlington

```{r}
tractcentroids <- st_centroid(tractstx)
tractsarlington <- tractstx[st_within(tractcentroids, arlington, sparse = FALSE), ]
arlingtontracts <- tractsarlington$GEOID

```

# Filter WAC in Arlington Tracts


# Sum by year and over all tracts to get citywide total and

```{r}
# Sum across all tracts and years
wac_arlington_df <- wac %>%
  filter(w_tract %in% arlingtontracts) %>%
  select(year, all_of(industryvars)) %>%
  group_by(year) %>%
  summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = -year, names_to = "industry_code", values_to = "job_count") %>%
  left_join(industry_labels, by = "industry_code") %>%
  select(Industry Name, year, job_count) %>%
  pivot_wider(names_from = year, values_from = job_count)
```

# Sort data Descending Order 
```{r}
wac_arlington_summary <- wac_arlington_df %>% arrange(desc(`2022`))
```

# 

```{r}
wac_arlington_change <- wac_arlington_summary %>%
  mutate(
    AbsoluteChange_2002_2012 = `2012` - `2002`,
    PercentageChange_2002_2012 = round(100 * (`2012` - `2002`) / `2002`, 1),
    AbsoluteChange_2002_2022 = `2022` - `2002`,
    PercentageChange_2002_2022 = round(100 * (`2022` - `2002`) / `2002`, 1)
  )

```


```{r}
readr::write_csv(wac_arlington_change, "ArlingtonTexasLEHD_2002_2022.csv")
```


```{r}
# tracts_jobs <- tractstx %>%
#   filter(GEOID %in% arlingtontracts) %>%
#   left_join(wac %>% select(w_tract, C000), by = c("GEOID" = "w_tract"))
# 
# ggplot() +
#   geom_sf(data = tracts_jobs, aes(fill = C000), color = "white", size = 0.3) +
#   geom_sf(data = arlington, fill = NA, color = "red", size = 1) +
#   scale_fill_viridis_c(option = "plasma", trans = "log10", labels = comma) +
#   labs(
#     title = "Job Distribution in Arlington, TX (2020)",
#     subtitle = "Total Jobs per Tract (LEHD WAC - JT00)",
#     fill = "Total Jobs",
#     caption = "Data: U.S. Census LEHD LODES (WAC), TIGER/Line (tracts)"
#   ) +
#   theme_minimal()
```

